---
title: "Assignment 5-Problem 1"
author: "Laura Ingulsrud"
date: "11/28/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r,echo = FALSE, message = FALSE}
library(tidyverse)
```
1) Male and female graduate enrollment (1967 - 2015). Compare trends in total graduate enrollment
for males and females (including full-time/part-time and private/public universities) in the United States from 1967 - 2015. Describe your results statistically, graphically and in text.https://github.com/lauraing/Assignment-5-Problem-1.git



```{r}

GradEnroll<- read.csv("Grad Enrollment 1967-2015.csv")


```


1) Male and female graduate enrollment (1967 - 2015). Compare trends in total graduate enrollment for males and females (including full-time/part-time and private/public universities) in the United States from 1967 - 2015. Describe your results statistically, graphically and in text.

```{r}

# compare graphically
Grad_total<- ggplot(GradEnroll, aes(x= Year))+
  geom_line(aes(y=Femails, color="Female"))+
  geom_line(aes(y=Males, color="Male"))+
  ylab("Enrolled Graduate Students")+
  ggtitle("Total US Male and Female Graduate Student (1967 - 2015) ")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(colour = "black")) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  scale_color_discrete(name = " ")
Grad_total


Grad_parttime<- ggplot(GradEnroll, aes(x= Year))+
  geom_line(aes(y=Fparttime, color="Female"))+
  geom_line(aes(y=Mparttime, color="Male"))+
  ylab("Enrolled Graduate Students")+
  ggtitle("US Graduate Student Male and Female Part Time Enrollment (1967 - 2015) ")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(colour = "black")) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
 scale_color_discrete(name = " ")


Grad_parttime

Grad_fulltime<- ggplot(GradEnroll, aes(x= Year))+
  geom_line(aes(y=Ffulltime, color="Female"))+
  geom_line(aes(y=Mfulltime, color="Male"))+
  ylab("Enrolled Graduate Students")+
  ggtitle("US Graduate Student Male and Female Part Time Enrollment (1967 - 2015) ")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(colour = "black")) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  scale_color_discrete(name = " ")


Grad_fulltime

Grad_Private_Public<-ggplot(GradEnroll, aes(x= Year))+
  geom_line(aes(y=private, color= "Private"))+
  geom_line(aes(y=public, color="Public"))+
  ylab("Enrolled Graduate Students")+
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.line = element_line(colour = "black")) +
  ggtitle("US Graduate Student Enrolled in Private and Public Schools (1967 - 2015)")+
  scale_color_discrete(name = " ")


Grad_Private_Public




```



```{r}
#  wilcoxon rank  

#Ho = Ranks are equal
#Ha= Ranks are not equal

wilcox.test(GradEnroll$Mfulltime, GradEnroll$Ffulltime)

#p value = .04 thus  the Female and male fulltime students differ significantly 

wilcox.test(GradEnroll$Mparttime, GradEnroll$Fparttime)
# p value <.001 thus the Female and male part time students differ significantly

wilcox.test(GradEnroll$private, GradEnroll$public)
# p value < .001 thus the number of students enrolled in public and private universities differ significantly

```